In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-24 paper/supplementary_main.tex
Updates from Overleaf

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{(v)}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$.

@@

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
\hat{c}|_{\Lambda^i}$, and let $\mathcal{E}(y)$ denote the set of
(one, two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
impose continuity of the concentrations, we set that the
concentrations when viewed from each edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure mass conservation, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n^i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n^i(y_{\rm in}) = 1, \quad
  n^i(y_{\rm{out}}) = -1, \quad
  n^i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
  \label{eq:def:sign}
\end{equation}

@@

\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

@@

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in
space to accurately capture sharp boundary layers.

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

@@

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ and $\bm \tau$ are the unit outward normal and unit tangent vectors to each facet, respectively. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space but not
on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/time_refinement_par_csf_mean.png}
        \caption{Mean tracer concentration in CSF and parenchyma under time step refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1\linewidth]{figures/time_refinement_art_ven_mean.png}
         \caption{Mean tracer concentration in PVSs around (veins and) arteries under time step refinement}
    \end{subfigure}
    \caption{Mean tracer concentrations after up to 24\,h in the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps of 1, 2, and 4 minutes (dt of 60, 120, or 240 seconds)}
    \label{fig:time_convergence_concentrations}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{figures/modelAMassConservation_total_conc.png}
    \caption{Total tracer content in the CSF, parenchyma, and arterial and venous PVS for a variant of the baseline model without tracer outflow. The total amount of tracer is constant after the initial influx phase demonstrating that the numerical scheme conserves mass.}
    \label{fig:mass_conservation}
\end{figure}

>>>>

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{a}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{v}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$. \mk{We always keep $R^i_2$, $R^i_1$ fixed in time in A.1.1 and A.1.4, right?}

@@

To complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{N} \subset \mathcal{V}$, write $\hat{c}^i =
\hat{c}|_{\Lambda^i}$, and let $\mathcal{E}(y)$ denote the set of
(one, two or three) edges in $\{\Lambda^i\}$ sharing the node $y$. To
impose continuity of the concentrations, we set that the
concentrations when viewed from each edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}^j(y)
  \qquad \foralls \Lambda^i, \Lambda^j \in \mathcal{E}(y)
  \qquad \foralls y \in \mathcal{B} .
  \label{eq:conc_continuity}
\end{equation}
Moreover, to ensure mass conservation, we set that the flux
going in and out at nodes should add to zero, or more precisely, that
for each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{\Lambda^i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c}^i(y) \right ) n^i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n^i : \mathcal{V} \rightarrow \{-1, \mk{0}, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n^i(y_{\rm in}) = 1, \quad
  n^i(y_{\rm{out}}) = -1, \quad
  \mk{n^i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} }.  
  \label{eq:def:sign}
\end{equation}
\mk{Why do we need the normal on non-IO vertices?}

@@

\subsubsection{Modelling CSF flow via the incompressible Stokes equations}
\label{sec:app:csf_flow}
The time-independent, incompressible Stokes equations model the flow
of an incompressible Newtonian fluid at low Reynolds numbers, and read
as follows: over the domain $\Omega_{\rm CSF} \subset \R^3$, find the
velocity vector field $\bm u : \Omega_{\rm CSF} \rightarrow \R^3$ and
the pressure field $p : \Omega_{\rm CSF} \rightarrow \R$ such that
\begin{subequations}
  \begin{alignat}{2}
    - \mu \Delta \bm u  + \nabla p &=  0 \quad && \text{ in } \Omega_{\rm CSF},
    \label{eq:momentum_equation} \\ 
    \nabla \cdot \bm u &= 0 \quad && \text{ in } \Omega_{\rm CSF}
    \label{eq:divergence_equation} .
  \end{alignat}
  \label{eq:stokes}%
\end{subequations}%
In addition, we impose the following boundary conditions to model flow
induced by CSF production in the choroid plexus with $\Gamma_{\rm AM-U}$ as the main efflux site:
\begin{subequations}
  \begin{alignat}{2}
    \mu \nabla \bm u \cdot \bm{n} - p \bm n &= -R_0 ( \bm u \cdot \bm n ) \bm n
    && \text{ on } \Gamma_{\mathrm{AM-U}}, \\
    \bm u &= 0 && \text{ on } \Gamma_{\mathrm{AM-L}}\cup\Gamma_{\rm{pia}}\cup\Gamma_{\rm{SSAS}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{pia}}, \\
    %\bm u &= 0 && \text{ on } \Gamma_{\rm{SSAS}}, \\
    \bm u \cdot \bm n & = \frac{1}{ |\Gamma_{\rm LV}|}  u_{\rm in}, \quad \bm u \cdot \bm \tau = 0 \quad && \text{ on } \Gamma_{\rm{LV}} ,  
  \end{alignat}
  \label{eq:stokes_bcs}%
\end{subequations}%
where $\bm n$ denotes the unit outward normal to the boundary, $\mu$ is the (dynamic) CSF viscosity, $R_0 > 0$ is a resistance parameter for CSF efflux, and $u_{\rm in}$ is a given fluid influx, here across the lateral ventricle wall $\Gamma_{\rm LV}$. We consider alternative variations of these boundary conditions in connection with estimating dispersion effects induced by CSF pulsatility, see~\Cref{sec:app:dispersion}.

@@

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in
space to accurately capture sharp boundary layers. \mk{Comment on Peclet number}

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

@@

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ is the unit outward normal vector to each facet. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space but not
on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/time_refinement_par_csf_mean.png}
        \caption{Mean tracer concentration in CSF and parenchyma under time step refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1\linewidth]{figures/time_refinement_art_ven_mean.png}
         \caption{Mean tracer concentration in PVSs around (veins and) arteries under time step refinement}
    \end{subfigure}
    \caption{Mean tracer concentrations after up to 24\,h in the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the standard resolution mesh for timesteps of 1, 2, and 4 minutes (dt of 60, 120, or 240 seconds)}
    \label{fig:time_convergence_concentrations}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{figures/modelAMassConservation_total_conc.png}
    \caption{Total tracer content in the CSF, parenchyma, and arterial and venous PVS for a variant of the baseline model without tracer outflow. The total amount of tracer is constant after the initial influx phase demonstrating that the numerical scheme conserves mass globally.}
    \label{fig:mass_conservation}
\end{figure}
